curatedMetagenomicData

### 2020-01-22 manuscript_NS_MRD.Rmd
updated

#drops unused levels
c.mnd.atlas <- droplevels(c.mnd.atlas,c("ISL","HUN","EST", "SVK", "LUX", "CAN", "AUT", "BGD","FJI","CHN","MDG","MNG","PER","RUS","TZA") )

@@

  p2 <<- p2 + stat_compare_means(comparisons = mnd.pairs) + xlab("MND Prevalence") + theme(legend.position="none")

@@

  p3 <- p2 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "NMDS",color = "MND Prevalence") + theme(legend.position = "none")
  #plot(p1 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence"))

@@

    p5 <- p4 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "MDS",color = "MND Prevalence") + theme(legend.position = "none")
    save_plot(paste("Beta_diversity_",title_text,"_MDS.png", sep = "" ), p5)

   #PCoA for compositional data with Bray-Curtis distances
    #   p4 <- plot_landscape(pseq.rel, method = "NMDS", distance = "bray", colour = "mnd", size = 3) 
    #   p4 <- p4 + scale_color_brewer(palette = "Dark2")+ scale_fill_gradient(low = "#e0ecf4", high = "#6e016b") 
    #p6 <- plot_landscape(pseq.subset, method = "MDS", distance = "bray", col = "mnd", size = 3) 
  #   # PCoA for compositional data with Bray-Curtis distances
   #p6 <- plot_landscape(microbiome::transform(pseq.core, "compositional"),
    #                    method = "PCoA", distance = "bray") +
    #     labs(title = paste("PCoA / Compositional / Bray-Curtis"))
  # 
    tsne_res <- tsne_phyloseq(pseq.subset, distance='bray', perplexity = 5, verbose=1, rng_seed = 3901)
    p7 <- plot_tsne_phyloseq(pseq.subset, tsne_res,color = 'mnd')
    p8 <- p7 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "TSNE",color = "MND Prevalence") + theme(legend.position = "bottom")
    save_plot(paste("Beta_diversity_",title_text,"_MDS.png", sep = "" ), p8)
       #this just generates the titles for the output
    title <- ggdraw() + 
    draw_label(
    title_text,
    fontface = 'bold',
    x = 0,
    hjust = 0
    ) +
    theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
    )

@@

# This calcualtes the group PERMANOVA perma_d(p.low,p.high,"TESTHigh_vesus_Low")
perma_d <- function(x,y,title_text) { # x & y must be either p.low,p.med or p.high
  # Convert to compositional data perma_d(p.low,p.high,"final_highVlow")
  pseqLvH <- merge_phyloseq(x,y) # pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p <<- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = FALSE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal(base_size = 12)  + annotate("text", x = 2.5, y = 0.01, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Beneficial Taxa") + scale_fill_discrete(name = "Taxa") + theme(axis.text.y = element_text(face="italic"))
  save_plot(paste("PERMANOVA_",title_text,".png", sep = "" ), p)
  #ggsave("test_permanova.png", device = "png",width = 297, height = 210, units = "mm")
 }

@@

pseq.subset <- subset_samples(pseq.c, !is.na(mnd) & disease == "healthy" & age_category != "child" & age_category != "newborn" & age_category != "schoolage" )

@@

beta_d(pseq.subset,"final_2")

@@

plot_abundance.ak = function(pseq.subset,title = "",
                         Facet = "Genus", Color = "Phylum"){
 # Arbitrary subset, based on Phylum, for plotting
 p1f = subset_taxa(pseq.subset, Genus %in% c("Akkermansia"))
 mphyseq = psmelt(p1f)
 mphyseq <- subset(mphyseq, Abundance > 0)
 p <- ggplot(data = mphyseq, mapping = aes_string(x = "mnd",y = "Abundance")) +
   geom_violin(fill = NA) +
   geom_point(size = 1, alpha = 0.3,
              position = position_jitter(width = 0.3)) + scale_y_log10()+
   theme(legend.position="none") + scale_colour_manual(values = wes_palette(glob_theme)) 
 library(rms)
fita <- lrm(mnd ~ gender + log(Abundance) + bmi_group + age_category, data =mphyseq)
fita <- lrm(mnd ~ gender + log(Abundance) , data =mphyseq)
fita <- lrm(mnd ~ gender + log(Abundance) + bmi_group, data =mphyseq)

save_plot(paste("akkermansia",title_text,".png", sep = "" ), p)
}

>>>>

#drops unused levels
c.mnd.atlas <- droplevels(c.mnd.atlas,c("ISL","HUN","EST", "SVK", "LUX", "CAN", "AUT", "BGD","FJI","CHN","MDG","MNG","PER","RUS","TZA", "KAZ"))

c.mnd.atlas <- subset_samples(c.mnd.atlas, !is.na(nationality) )

@@

  p2 <<- p2 + stat_compare_means(comparisons = mnd.pairs) + xlab("MND DALYs") + theme(legend.position="none")

@@

  p3 <- p2 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "NMDS",color = "MND DALYs") + theme(legend.position = "none")
  #plot(p1 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence"))

@@

    p5 <- p4 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "MDS",color = "MND DALYs") + theme(legend.position = "none")
    save_plot(paste("Beta_diversity_",title_text,"_MDS.png", sep = "" ), p5)

   #PCoA for compositional data with Bray-Curtis distances
    #   p4 <- plot_landscape(pseq.rel, method = "NMDS", distance = "bray", colour = "mnd", size = 3) 
    #   p4 <- p4 + scale_color_brewer(palette = "Dark2")+ scale_fill_gradient(low = "#e0ecf4", high = "#6e016b") 
    #p6 <- plot_landscape(pseq.subset, method = "MDS", distance = "bray", col = "mnd", size = 3) 
  #   # PCoA for compositional data with Bray-Curtis distances
   #p6 <- plot_landscape(microbiome::transform(pseq.core, "compositional"),
    #                    method = "PCoA", distance = "bray") +
    #     labs(title = paste("PCoA / Compositional / Bray-Curtis"))
  # 
    tsne_res <- tsne_phyloseq(pseq.subset, distance='bray', perplexity = 5, verbose=1, rng_seed = 3901)
    p7 <- plot_tsne_phyloseq(pseq.subset, tsne_res,color = 'mnd')
    p8 <- p7 + scale_colour_manual(values = wes_palette(glob_theme)) + labs(title = "TSNE",color = "MND DALYs") + theme(legend.position = "bottom")
    save_plot(paste("Beta_diversity_",title_text,"_MDS.png", sep = "" ), p8)
       #this just generates the titles for the output
    title <- ggdraw() + 
    draw_label(
    title_text,
    fontface = 'bold',
    x = 0,
    hjust = 0
    ) +
    theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
    )

@@

# This calcualtes the group PERMANOVA perma_d(p.low,p.high,"TESTHigh_vesus_Low")
perma_d <- function(x,y,title_text) { # x & y must be either p.low,p.med or p.high
  # Convert to compositional data perma_d(p.low,p.high,"final_highVlow")
  pseqLvH <- merge_phyloseq(x,y) # pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p <<- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = FALSE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal(base_size = 12)  + annotate("text", x = 2.5, y = 0.01, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Beneficial Taxa") + scale_fill_discrete(name = "Taxa") + theme(axis.text.y = element_text(face="italic")) + theme_void()
  save_plot(paste("PERMANOVA_",title_text,".png", sep = "" ), p)
  #ggsave("test_permanova.png", device = "png",width = 297, height = 210, units = "mm")
 }

@@

pseq.subset <- subset_samples(pseq.c, !is.na(mnd) & is.na(disease) & disease == "healthy" & age_category != "child" & age_category != "newborn" & age_category != "schoolage" )

@@

beta_d(pseq.subset,"final_2")
beta_d(atlas.pseq,"final_3")

@@

plot_abundance.ak = function(pseq.subset,title = "",
                         Facet = "Genus", Color = "Phylum"){
 # Arbitrary subset, based on Phylum, for plotting
 p1f = subset_taxa(pseq.subset, Genus %in% c("Akkermansia"))
 mphyseq = psmelt(p1f)
 mphyseq <- subset(mphyseq, Abundance > 0)
 p <- ggplot(data = mphyseq, mapping = aes_string(x = "mnd",y = "Abundance")) +
   geom_violin(fill = NA) +
   geom_point(size = 1, alpha = 0.3,
              position = position_jitter(width = 0.3)) + scale_y_log10()+
   theme(legend.position="none") + scale_colour_manual(values = wes_palette(glob_theme)) 
 #library(rms)
#fita <- lrm(mnd ~ gender + log(Abundance) + bmi_group + age_category, data =mphyseq)
#fita <- lrm(mnd ~ gender + log(Abundance) , data =mphyseq)
#fita <- lrm(mnd ~ gender + log(Abundance) + bmi_group, data =mphyseq)

save_plot(paste("akkermansia_",title_text,".png", sep = "" ), p)
}
